Symbolic linearization of equations of motion of constrained multibody systems

### 2013-08-02 PetersonGedeHubbard2012.tex
Wordsmithed introduction.

\begin{abstract}
% Context
Many common dynamic mechanical systems have configuration or velocity constraints.
% Need  (what we have and what we want)
Analyses of such systems often require linearized forms of the motion
equations.
% Task
To address this issue, we developed a procedure for organizing the constraint
and motion equations and their subsequent linearization.
This procedure was specifically developed for equations of motion generated by
Kane's method, but is compatible with any method as long as the following are
true: only ordinary differential equations are required to describe the system,
and independent and dependent states remain in the final equations (i.e. no
DAEs, no substitution of variables).
% Object of the document
Following a brief review of Kane's method and the structure of the equations it
generates, we present the procedure to symbolically linearize the nonlinear
motion equations of constrained multibody systems, and illustrate it with the
% TODO: change name of example
familiar example of the rolling disc.

@@

\section{Introduction}
\label{sec:intro}
Designing a feedback controller or performing stability analysis are two common
analyses performed on models of multibody systems.
A linear form of the equations of motion is required to apply the most
frequently used techniques for the analyses to such systems.
Normally, obtaining the linearized equations of motion is as simple as
calculating the Jacobian of the right hand side function of the equations of
motion with respect to the states.
However, when the multibody system in questions has constraints, this process
is more complicated.
If one has used Kane's method \cite{Kane1985} to generate the equations for a
constrained system, the resulting equations will have a particular form: a set
of ordinary differential equations containing both the independent and
dependent states.
Simple inspection of the equations does not reveal that they represent a
constrained system - their form is similar to the equations of an unconstrained
system.
It is critical to understand linearizing these equations by computing the
Jacobian gives incorrect results.
This paper provides a way to correctly generate the linear equations of motion
for constrained systems which have been found using Kane's method (or any other
method which results in equations with the same form).

Multibody systems (rigid or flexible) and particles often have constraints
which limit how they can move or be configured.
In this paper, we will discuss three types of constraints: configuration,
velocity, and acceleration.
Configuration constraints limit the location or orientation of parts of the
system, relative to the external world or other parts of the system.
Velocity constraints limit the speeds at which the configuration can change,
either from configuration constraints (through time differentiation) or
independent application.
Velocity constraints most often appear in systems where there is rolling
without slip or there are closed kinematic loops.
Within this paper, acceleration constraints refer to time differentiated
velocity constraints.

% TODO: Improve literature review.
Bottema appears to be the first author to have shown that special
considerations are needed for linearizing nonholonomic systems
\cite{Bottema1949}.
While other authors have examined linearization of nonholonomic systems, we
have found isssues with applying any of them to equations generated by Kane's
method.
Kang et al. \cite{Kang2003} and Negrut and Ortiz \cite{Negrut2006} have both
explored linearization, but these methods have been developed for equations of
motion found using Lagrange's method, which contain Lagrange multipliers (which
are not present in equations of motion generated from Kane's method).
Minaker and Rieveley \cite{Minaker2010} and Schwab and Meijaard
\cite{Schwab2003} both have developed techniques for generating equations of
motion (and linear forms thereof), however both techniques are more numerical
in nature - they do not provide symbolic equations of motion or symbolic
linearized equations.
These processes, along with others such as Neuman's %TODO: fix reference
, formulate the linear equations as a step in the equations of motion
generation process.
They are not capable of linearizing existing equations of motion.

% TODO:reiterate limiations of model
The goal of this paper is to correctly establish the first order relationships
(about an arbitrary point of linearization) between a selection of independent
coordinates and speeds and the time-derivatives of all coordinates and speeds
(both independent and dependent). The formalism presented here is generic
enough to cover most examples of time-varying constrained multibody systems
with arbitrary external inputs and arbitrary specified quantities.
% TODO: electronic supplementary material with examples for other methods
% TODO: rewrite following sentences after e.s.m.
The
procedure was created for systems whose equations of motion have been derived
with Kane's method. Although also applicable to dynamic system equations
formulated using other methods, it is restricted to systems which can be
completely described by a set of ODEs (i.e., DAEs needn't be solved).

% TODO:rewrite organization of paper
The use of Kane's method in generating equations of motion is reviewed first,
in order to properly orient the reader to the format and some of the qualities
of the generated equations.
Kane's method is then used to generate the equations of motion for a simple
system,
%TODO: update example name
a rolling disk with a trailing arm.
The need for a different linearization technique is also demonstrated with this
example.
In response to this need, we present the newly developed linearization
procedure and its derivation.
This new linearization procedure is then performed on the system to provide an
example of its application.
Finally, we discuss other details and nuances of the procedure and its use.

>>>>

\begin{abstract}
% Context
Many common dynamic mechanical systems have configuration or velocity constraints.
% Need  (what we have and what we want)
Analyses of such systems often require linearized forms of the motion
equations.
% Task
To address this issue, we developed a procedure for organizing the constraint
and motion equations and their subsequent linearization.  This procedure was
specifically developed for equations of motion generated by Kane's method, but
is compatible with any method as long as the following are true: only ordinary
differential equations are required to describe the system, and independent and
dependent states remain in the final equations (i.e. no DAEs, no substitution
of variables).
% Object of the document
Following a brief review of Kane's method and the structure of the equations it
generates, we present the procedure to symbolically linearize the nonlinear
motion equations of constrained multibody systems, and illustrate it with the
familiar example of the rolling disc.

@@

\section{Introduction}
\label{sec:intro}
Design of linear state feedback controllers and first order stability analysis
of equilibria are two of the most common uses of multibody system dynamics
models. Both of these uses require linearized equations of motion. For
unconstrained systems, obtaining the linearized equations of motion is as
simple as arranging the equations of motion in first order form ($\dot{x} =
f(x, t)$) and evaluating the Jacobian of the right hand side function, with
respect to the states, evaluated at the equilibria point of interest.  However,
when the system in questions has constraints, i.e., not all the states are
independent, this process is not as straightforward.

For a system system with constraints, Kane's dynamical equations\cite{Kane195}
are first ordinary differential equations (in the generalized speeds) which
include both the independent and dependent states (assuming the analyst has not
taken steps to algebraically eliminate states deemed to be dependent). The
dynamical equations and the kinematical differential equations can be solved
for the time derivatives of the generalized coordinates and generalized speeds,
respectively. Taken together, it is tempting to think that this system of first
order differential equations could be linearized in the same fashion as an
unconstrained system. Unfortunately, this is not the case. Linearizing these
equations in the same manner as one would for an unconstrained system (by
directly computing the Jacobian) is incorrect and will yield incorrect
linearized equations (and subsequently incorrect eigenvalues, in the case of a
stability analysis). We illustrate this by means of a familiar example, the
rolling disc, formulated with a non-traditional set of coordinates and
generalized speeds which are not all independent. We then provide a systematic
method to correctly generate the linear equations of motion for constrained
systems which have been found using Kane's method (or any other method which
results in equations with the same form). We then apply this method to the
rolling disc example to illustrate that it yields eigenvalues that match
previously publish results.

We consider systems subject to three types of constraints: configuration,
velocity, and acceleration.  Configuration constraints limit the location or
orientation of parts of the system, relative to the external world or other
parts of the system.  Velocity constraints limit the speeds at which the
configuration can change, either from configuration constraints (through time
differentiation) or independent application.  Velocity constraints most often
appear in systems where there is rolling without slip or there are closed
kinematic loops.  In the context of this paper, acceleration constraints arise
either from time differentiated velocity constraints or similar kinematic
considerations.

% TODO: Improve literature review.
Bottema appears to be the first author to have shown that special
considerations are needed for linearizing nonholonomic systems
\cite{Bottema1949}. While other authors have examined linearization of
nonholonomic systems, we have found that most are not directly applicable to
equations formed using Kane's method, and while we do not doubt that they
\textit{could} be applied correctly, the details of how to do so, as far as the
authors are aware, have not been explicitly presented. Kang et al.
\cite{Kang2003} and Negrut and Ortiz \cite{Negrut2006} have both explored
linearization, but these methods have been developed for equations of motion
found using Lagrange's method, which contain Lagrange multipliers (which are
not present in equations of motion generated from Kane's method). Further, both
of these papers lacked a thorough discussion of basic concepts such as how many
quantities are independent, and in the case of eigenvalue computation, how many
eigenvalues should be present. Minaker and Rieveley \cite{Minaker2010} and
Schwab and Meijaard \cite{Schwab2003} both have developed techniques for
generating equations of motion (and linear forms thereof), however both
techniques are numerical in nature - they do not provide symbolic equations of
motion or symbolic linearized equations. While Neuman presents a symbolic (as
opposed to numeric) formulation for linearization of the Q-matrix formulation
of the Lagrangian\cite{Neuman1984}, his work makes no mention of constrained
systems and attempts to contact the author regarding the software (algebraic
robot modeler) as well as internet searches for the software have been
fruitless.

% TODO:reiterate limitations of model
The goal of this paper is to definitively establish how the first time
derivative of all coordinates and speeds (independent and dependent) depend, to
first order, upon a selection of independent coordinates and independent
speeds, for an arbitrary point of linearization. We demonstrate how a reduced
portion of these linearized equations can be used to perform standard stability
analysis (i.e., for control system design or linear stability analysis). The
formalism presented here is generic enough to cover most examples of
time-varying constrained multibody systems with arbitrary external inputs and
arbitrary specified quantities.
% TODO: electronic supplementary material with examples for other methods
% TODO: rewrite following sentences after e.s.m.
The procedure was created for systems whose equations of motion have been
derived with Kane's method. Although also applicable to dynamic system
equations formulated using other methods, it is restricted to systems which can
be completely described by a set of ODEs (i.e., DAEs needn't be solved).

We begin with a review of Kane's method for generating equations of motion is
reviewed first, in order to properly orient the reader to the format and some
of the qualities of the generated equations. Next, we apply Kane's method to
generate the equations of motion for the rolling disk, but we purposefully
introduce a dependent coordinate and three dependent speeds in the formulation.
The need for a different linearization technique is demonstrated with this
example. In response to this need, we present the newly developed linearization
procedure and its derivation.  This new linearization procedure is then applied
the same rolling disk example to illustrate its validity and an example of its
application.  Finally, we discuss other details and nuances of the procedure
and its use, as well as directions for future work.
